---
title: "A First Simulation Example on Designing and Assessing a Regression Function"
author: "Victoire Djimna Noyum"
date: "20/05/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the packages

```{r, warning=FALSE, message=FALSE}
library(mvtnorm)
library(tidyverse)
```

## First Simulation Example on Designing and Assessing a Regression Function

## binormal distribution : mean = 0 and rho = 0.8

```{r}
set.seed(8725)
sigma = matrix(c(1,0.8, 0.8,1),2,2)    # covariance matrix
sim10 <- rmvnorm(n=10, mean = c(0,0),sigma = sigma)
sim10 <- data.frame(X = sim10[,1], Y = sim10[,2])
sim10
```


## Fit the data to a linear model
```{r}
model1 <- lm(Y~X,data = sim10)
summary(model1)
```

###  Mean Squared Error (MSE)

```{r}
MSE <- sum(model1$residuals^2)/10
MSE
```


##  Plot the linear model, data and best regression function

```{r, warning=FALSE, message=FALSE}
p1 <- ggplot(data = sim10,aes(x = X, y = Y))+
  geom_point()+
  stat_smooth(method = lm)+
  geom_abline(intercept = 0, slope = sqrt(0.8), color="red",
              linetype="dashed", size=1)
p1
```

## Generate the large data (1000 observations)

```{r}
set.seed(8725)
sigma = matrix(c(1,0.8, 0.8,1),2,2)
sim1000 <- rmvnorm(n=1000, mean = c(0,0),sigma = sigma)
sim1000 <- data.frame(X = sim1000[,1], Y = sim1000[,2])
sim1000
```

### True error rate

```{r}
model2 <- lm(Y~X,data = sim1000)
summary(model2)

pred <- predict(model1, newdata = sim1000)

test <- data.frame(actual=sim1000$Y, pred = pred)
test$error <- with(test, pred-actual)


errtr <- with(test, mean(error^2))
errtr

```


### Perfomance of best regression function: err*

```{r}
errstar <- sum(model2$residuals^2)/1000
errstar
```


## Monte-Carlo (MC) : simulation by repeating the above for 500 training sets of the same size (10 observations) and the same large testing sets (1000 observations)

### Plot showing the best regresson function (in bold) and the 500 linear models (in gray).

```{r, message=FALSE, warning=FALSE}
simfunct <- function(v){
  sigma = matrix(c(1,0.8, 0.8,1),2,2)
  sim <- rmvnorm(n=v, mean = c(0,0),sigma = sigma)
  sim <- data.frame(X = sim[,1], Y = sim[,2])
  return(sim)
} 
p2 <- ggplot()+
  theme_bw()

Vect_MSE <- NULL
Err <- NULL
for (i in 1:500) {
w <-   simfunct(20)
mod1 <- lm(Y~X,data = w)
MSE1 <- sum(mod1$residuals^2)/10
Vect_MSE <- c(Vect_MSE, MSE1)
p2 <- p2+stat_smooth(method = lm,se = FALSE,data = w,col ="gray",aes(x=X, y=Y))
 
pred1 <- predict(mod1, newdata = sim1000)

test <- data.frame(actual=sim1000$Y, pred = pred1)
test$error <- with(test, pred1-actual)

errtr <- with(test, mean(error^2))
Err <- c(Err,errtr)
}

p2+geom_abline(intercept = 0, slope = sqrt(0.8), color="black",
                     linetype="dashed", size=1.2)

```


### 500 MSE values

```{r}
Vect_MSE
```

### Mean of true errors rate

```{r}
meanv <- mean(Err)
meanv 
```

### Variance of true errors rate

```{r}
varerr <-  sum((Err-meanv)^2)/(499)
varerr
```


## Repetition of the obove in the following training-set sizes:  20, 40, 80, 100, 200, 300, 400, 500, 700, 1000


### Plot of the performance of best regression function, the mean and variance of true errors rate.  

```{r, warning=FALSE, message=FALSE}
simulfunct <- function(vect){
  sigma = matrix(c(1,0.8, 0.8,1),2,2)
  sim1000 <- rmvnorm(n=1000, mean = c(0,0),sigma = sigma)
  sim1000 <- data.frame(X = sim1000[,1], Y = sim1000[,2])
  
  vect_errstar <- NULL
  vect_mean <- NULL
  vect_var <- NULL
  for (ntr in vect) {
    model <- lm(Y~X,data = sim1000)
    errstar <- sum(model$residuals^2)/1000
    vect_errstar <- c(vect_errstar,errstar)
    Vect_MSE <- NULL
    Err <- NULL
    meanv <- NULL
    for (i in 1:500) {
      sim <- rmvnorm(n=ntr, mean = c(0,0),sigma = sigma)
      sim <- data.frame(X = sim[,1], Y = sim[,2])
      mod1 <- lm(Y~X,data = sim)
      pred1 <- predict(mod1, newdata = sim1000)
      test <- data.frame(actual=sim1000$Y, pred = pred1)
      test$error <- with(test, pred1-actual)
      errtr <- with(test, mean(error^2))
      Err <- c(Err,errtr)
    }
    meanv <- mean(Err)
    vect_mean <- c(vect_mean, meanv)
    
    
    varerr <-  sum((Err-meanv)^2)/(499)
    vect_var <- c(vect_var, varerr)  
  } 
  dt <- data.frame(errs_tar=vect_errstar, mean_tr= vect_mean, var_tr= vect_var )
  return(dt)
}

ve <- c(20, 40, 80, 100, 200, 300, 400, 500, 700, 1000)

dat <- simulfunct(ve)
dat$ve <- ve
p <- ggplot(data=dat,aes(x = ve))+
  geom_line(aes(y=mean_tr,colour = "mean_tr"),size = 1)+
  geom_line(aes(y=var_tr, colour = "var_tr"), size = 1)+
  geom_line(aes(y = errs_tar, colour = "errs_tar"),size = 1)+
  theme_bw()+
  scale_colour_manual("", 
                      values = c("mean_tr"="green", "var_tr"="red", 
                                 "errs_tar"="blue"))+
  ylim(0,0.43)+
  ylab("")
p
```



